---
title: "Module 1: Dealing with Data"
output: 
  html_notebook: 
    highlight: pygments
    theme: sandstone
    css: other.css
editor_options: 
  chunk_output_type: inline
---

```{r init, echo=FALSE}
knitr::opts_chunk$set(
  echo=T, 
  eval = F,
  message = F, 
  warning = F, 
  comment = NA,
  R.options=list(width=120), 
  cache.rebuild=F, 
  cache=T,
  fig.align='center', 
  fig.asp = .7,
  dev = 'svg', 
  dev.args=list(bg = 'transparent')
)

```

## Setup

What is the `tidyverse`?

The `tidyverse` consists of a few key packages for data import, manipulation, visualization and more.

```{r setup}
library(tidyverse)
```


## Objects and Classes

```{r objects, echo=FALSE}
x = 1:3
y = 'a'
z = list(one = x, two = y)

x
y
z
```

```{r inspect}
str(z)
class(y)
```

## Data Structures

Vectors form the basis of R data structures. Two main types are atomic and lists.

```{r vector}
my_vector <- c(1, 2, 3)   # standard vector
```

```{r list}
my_list <- list(a = 1, b = 2)   # a named list
my_list
```

## Data frames

Data frames are a special kind of list, and probably the most commonly used for data science purposes.

```{r data_frame, eval=TRUE}
my_data = data.frame(
  id = 1:3,
  name = c('Vernon', 'Ace', 'Cora')
)

my_data
class(my_data)
```

## Importing Data

Importing data is usually the first step.

```{r import, eval=TRUE}
demographics = read.csv('data/demos_anonymized.csv')
ids = read.csv('data/ids_anonymized.csv')
```

## Working with Databases

Databases must be connected to, but otherwise are used just like data frames.

```{r databases, eval=FALSE}
# requires DBI and RSQLite packages
library(DBI)
con <- dbConnect(RSQLite::SQLite(), ":memory:")
# con

copy_to(con, demographics, 'demos')
```



## Selecting Columns

A common step is to subset the data by column.

```{r select1}
demographics %>% 
  select(gender, age, libuser)
```


```{r select2}
demographics %>% 
  select(-libuser)
```

```{r select3}
demographics %>% 
  select(starts_with('award'))
```


## Filtering Rows

To filtering data, think of a logical statement, something that can be `TRUE` or `FALSE`.

```{r filter}
my_filtered_data = demographics %>% 
  filter(age < 40)

my_filtered_data = demographics %>% 
  filter(libuser == 1)
```


## Generating new data

Another very common data processing task is to generate new variables.

```{r mutate}
demographics = demographics %>% 
  mutate(new_age = (age - mean(age, na.rm = T))/sd(age, na.rm = T))   
```

## Renaming columns



```{r rename1}
demographics = demographics %>% 
  rename(age_std = new_age)
```

```{r rename2}
demographics %>% 
  rename_all(toupper) %>% 
  colnames()
```

## Merging

Merging data can take on a variety of forms, and depending on the data, can be be quite complicated.

```{r example_joins}
# same N rows as demos
left_join(demographics, ids)

# only ~ 50k rows
inner_join(demographics, ids) 
```

## Exercises

### Selecting and filtering

Use the `:` operator to select successive columns.

```{r ex1a, eval=FALSE}
colnames(demographics)

demographics %>% 
  select(?)
```

Filter the data to award amounts less than 500000.


```{r ex1b, eval=FALSE}
demographics %>% 
  filter(award_total_amount ?)
```

### Generating new data

Generate a new award amount variable that is the log of the original.  Give the new variable a useful name.

```{r ex2, eval=FALSE}
demographics %>% 
  mutate(? = log(?))
```


## Python examples

Using Python for data science is not far removed from R

Python's main data processing module is `pandas`


### Import

```{python py_import, engine.path= '/Users/micl/anaconda3/bin/python'}
import pandas as pd
import numpy  as np

demographics = pd.read_csv('data/demos_anonymized.csv')
ids = pd.read_csv('data/ids_anonymized.csv')

demographics.head()  # show a few lines
```

### Selecting Columns

```{python py_select}
# select by name
demographics[['age', 'award_total_amount']]
```


```{python py_select2}
# select successive columns
demographics.loc[:,'libuser':'age']
```


```{python py_select3}
# select by pattern
demographics.filter(regex='^award') 
```

### Filtering Rows

```{python py_filter}
my_filtered_data = demographics[demographics.libuser == 1]
my_filtered_data.libuser.nunique()
```

### Generating new data

```{python py_mutate}
demographics[['new_age']] = (demographics[['age']] - np.mean(demographics[['age']])) / np.std(demographics[['age']])

demographics.new_age.describe()  # mean = 0  sd = 1
```

### Renaming columns

```{python py_rename}
demographics = demographics %>% 
  rename(age_std = new_age)
```


### Joins

```{python py_left_join}
demos_joined = pd.merge(demographics, ids, how='left', on='EMPLID')
demos_joined
```

```{python py_left_join2}
demos_joined = demographics.join(ids, how='left', lsuffix='EMPLID')
demos_joined.shape
demos_joined.columns
```


```{python py_inner_join}
demos_joined = demographics.join(ids, how='inner', lsuffix='EMPLID')

demos_joined.columns
```

