---
title: "Module 2: Summarizing and Visualizing Data"
output: 
  html_notebook: 
    highlight: pygments
    theme: sandstone
    css: other.css
editor_options: 
  chunk_output_type: inline
---

```{r init, echo=FALSE}
# these options are primary useful to the creation of the html document
knitr::opts_chunk$set(
  echo=T, 
  eval = F,
  message = F, 
  warning = F, 
  comment = NA,
  R.options=list(width=120), 
  cache.rebuild=F, 
  cache=F,
  fig.align='center', 
  fig.asp = .7,
  dev = 'svg', 
  dev.args=list(bg = 'transparent')
)
```

```{r catchup, eval=TRUE}
library(tidyverse)
demographics = read.csv('data/demos_anonymized.csv')
ids = read.csv('data/ids_anonymized.csv')
model_variables = read.csv('data/model_variables_anonymized.csv')
```

## Introduction

## Get a Quick Summary

It's always a good idea to glance at the data

```{r summary}
glimpse(demographics)
```
```{r}
summary(demographics[,1:10])
```



## Grouping and Summarizing Operations

A very common operation is to do things by group(s) then create new summary variables.

```{r group_by, eval=TRUE}
demographics %>% 
  group_by(libuser) %>% 
  summarise(age = mean(age, na.rm = T)) 
```



## Grouping and Summarizing Operations

Do multiple operations at once

```{r group_by2, eval=TRUE}
demographics %>% 
  group_by(libuser) %>% 
  summarise(age_mean = mean(age, na.rm = T),
            age_sd = sd(age, na.rm = T),
            age_max = max(age, na.rm = T),
            prop_male = mean(gender=='Male', na.rm = T)) 
```



##  Mapping functions to groups

We’ll use the `map` function to *map* the `sum` function to each element in the list

```{r map, eval=TRUE}
x = list(1:3, 4:6, 7:9)
map(x, sum)
```



## Grouping and Summarizing Operations

We can even do this with modeling and other operations. 

This extracts the coefficients from a model run for each group.

```{r group_by3, eval=TRUE}
demographics %>% 
  drop_na(race) %>% 
  group_split(race) %>%   
  map(~lm(award_total_amount ~ gender, data = .))
```


## Visualization

Visualizing data requires that you:
- Consider carefully the information you want to display
- And then how you want to display it

Tell a story with the data.

And have some fun with it!


## ggplot2

The most widely used visualization package in R



## Layers

Visualization can be thought of in a layered fashion

Start with the base, then build up

## More pipes

`ggplot2` uses a `+` as a pipe to add layers

```{r ggplot2, eval=FALSE}
library(ggplot2)
ggplot(data) +
  geom_point(aes(x = var1, y = var2)) +
  geom_line() +
  theme(plot.caption = element_text(size = 6))
```

We can pipe *to* any ggplot as we did before (` %>% `)



## Aesthetics

**Aesthetics** (`aes`) map variables to visual properties

**Geoms** are the geometric units we want to display


## Aesthetics

```{r aesthetics}
model_variables %>% 
  filter(award_total_amount > 1e7) %>% 
  ggplot() +
  geom_density(aes(x = award_total_amount, 
                   color = factor(gender),
                   fill = factor(gender)),
               alpha = .2) + 
  scale_x_continuous(breaks = (1:10) * 1e7, trans = 'log')
```

```{r aesthetics2}
demographics %>% 
  filter(award_year_start < 2020 & award_year_start > 1990) %>%  
  ggplot(aes(award_year_start, award_total_log)) + 
  geom_smooth(aes(color=factor(libuser)))
```


## Stats

We can also use ggplot2 to create statistics we want to visualize

Typically used indirectly when geoms are called

Can be used for more direct control

```{r ggstats2}
ggplot(model_variables, aes(age, award_total_amount)) +
  geom_point(alpha = .02) +
  stat_ellipse(color = '#ff5500')
```



## Scales

Scales are used to add specifications to axes, colors, etc.

```{r scales}
model_variables %>% 
  filter(award_total_amount >= 1e6) %>% 
  ggplot() +
  geom_density(aes(x = award_total_amount, 
                   color = gender,
                   fill = gender),
               alpha = .2) + 
  scale_x_continuous(breaks = c(1e6, 5e6, 1e7, 2.5e7, 5e7), 
                     trans = 'log') +
  scale_fill_viridis_d(begin = .25, end = .5) +
  scale_color_viridis_d(begin = .25, end = .5) 
```


## Facets

Facets allow another dimension to plots by group

- **facet_grid**: returns a matrix of like dimensions
- **facet_wrap**: more flexible specification

```{r facets}
model_variables %>% 
  ggplot() +
  geom_density(aes(x = award_total_amount, 
                   color = libuser,
                   fill = libuser),
               alpha = .2) +
  facet_wrap(~gender)
```




## Themes

`Themes` allow for customization

Two uses of a 
- a built-in versions (e.g. `theme_minimal`)
- DIY  (`theme`(...))

For the theme function, each argument, takes on a specific value or an element function:

- `element_rect`
- `element_line`
- `element_text`
- `element_blank`


## Themes

```{r themes1}
model_variables %>% 
  ggplot() +
  geom_smooth(aes(x = age,
                y = award_total_amount,
                color = libuser),
               alpha = .2) + 
  theme_minimal()
```


```{r themes2}
model_variables %>% 
  ggplot() +
  geom_smooth(aes(x = age,
                y = award_total_amount,
                color = libuser),
               alpha = .2) + 
  theme(axis.text.x = element_text(size=12),
        panel.grid.minor.x = element_blank(),
        plot.background = element_rect(color = 'rosybrown'),
        panel.background = element_rect(fill = 'papayawhip'))
```

## Interactivity

Interactivity is a must-have tool for web-based presentation

Use to enhance exploration of the data
- Not just because one can

Allows for additional dimensions

Even useful for exploring raw data



## Interactivity


General 

- [plotly](https://plot.ly/r/)
  - also used in Python, Matlab, Julia
  - can convert `ggplot2` images to interactive ones
  
- [highcharter](http://jkunst.com/highcharter/)
  - general wrapper for highcharts.js 
  - works with some R packages out of the box
  
- [rbokeh](http://hafen.github.io/rbokeh/)
  - like plotly, it also has cross language support



## Interactivity

Specific functionality:

- [DT](https://rstudio.github.io/DT/)
  - interactive data tables
  
- [leaflet](https://rstudio.github.io/leaflet/)
    - maps with OpenStreetMap
    
- [visNetwork](http://datastorm-open.github.io/visNetwork/)
    - Network visualization
    


## Plotly

`traces`
- `add_`, work similar to geoms
  
`modes`
- allow for points, lines, text and combinations

`aesthetics`
- variables are denoted with `~`, constants do not use
- `x =~ var1` vs `x = 2`




## Plotly

Plotly uses the standard pipe `%>%`

```{r plotly}
library(plotly)

model_variables %>% 
  plot_ly(x = ~gender, y = ~ age) %>% 
  add_boxplot(color =~ gender)
```



## Plotly

```{r plotly1}
library(plotly)

model_variables %>% 
  group_by(libuser, gender) %>% 
  summarise(award = mean(award_total_amount)) %>% 
  plot_ly(x = ~libuser, 
          y = ~award, 
          color = ~gender,
          text = ~round(award), 
          textposition = 'auto', 
          type = 'bar') %>% 
  layout(bargap = 0.25, 
         bargroupgap = 0.25)
```



## Plotly

```{r plotly2}
init = glm(award_total_amount >= 5000000 ~ age*libuser, 
           data = model_variables, 
           family = binomial)

model_variables %>% 
  modelr::add_predictions(init, type = 'response') %>% 
  plot_ly(x = ~age, y = ~ pred) %>% 
  add_lines(color =~ libuser, line = list(shape = "spline")) %>% 
  layout(title = 'Predicted Prob. Award > 1 mil')
```



## Plotly

Use ggplotly to turn our formerly static plots into interactive ones. 

```{r plotly3}
p = model_variables %>% 
  ggplot() +
  geom_density(aes(x = log(award_total_amount), 
                   color = libuser,
                   fill = libuser),
               alpha = .2) + 
  facet_wrap(~ gender) 
ggplotly()
```




## Python examples

### Init

```{python py_init, engine.path= '/Users/micl/anaconda3/bin/python'}
import pandas as pd
import numpy as np

demographics = pd.read_csv('data/demos_anonymized.csv')
ids = pd.read_csv('data/ids_anonymized.csv')
model_variables = pd.read_csv('data/model_variables_anonymized.csv')
```


###  Grouping and Summarizing data

```{python py_summary}
demographics.describe(include = 'all')
```
```{python py_summary_numeric}
demographics.describe(include = [np.number])
```

```{python py_summary_string}
demographics.describe(include = [np.object])
```




```{python py_group_by, eval=TRUE}
lib_group = demographics.groupby('libuser', sort=True, )

# automatically chooses numeric
lib_group.mean()
```


```{python py_group_by2, eval=TRUE}
lib_group.get_group(0).head()
```


```{python py_group_by3, eval=TRUE}
lib_group.size()
```

```{python py_group_by4, eval=TRUE}
lib_group.describe()
```
### Mapping a function

```{python py_map}
x = [[1,2,3], [4,5,6], [7,8,9]]
list(map(np.sum, x))
```

```{python py_map2}
demographics.select_dtypes('number').apply(np.mean)
```


### Visualization

`matplotlib` is the most common visualization module in Python, though it's fairly dated at this point. As such we'll use a ggplot implementation in python called `plotnine`.

Unfortunately for plotly, the interactivity makes it unusable within the R notebook (at present), so you may need to switch to Anaconda or other IDE to try other modules like `plotly`.  Even Python users will still use R for easier visualization though, so feel free to do what you like there, then use ggplot etc. in R when the time comes.

That said, I'll show a couple plots

```{python plotnine}
import plotnine
```

```{python gg1}
dplot = demographics[(demographics.award_year_start < 2020) & (demographics.award_year_start > 1990) & (demographics.award_total_log > 11)]
dplot
ggplot(dplot, aes(x='award_year_start', y='award_total_log')) + geom_smooth(aes(group = 'libuser', color = 'libuser'))
```


```{python gg2}
year_award_average = demographics[demographics.libuser == 1].groupby(['award_home_dept', 'award_year_start']).mean()
year_award_average
```



```{python py_plotly_init}
import plotly
import plotly.plotly as py
from plotly.offline import init_notebook_mode
import plotly.graph_objs as go
plotly.offline.init_notebook_mode(connected=True)
```


```{python py_plotly}
y0 = np.random.randn(50)-1
y1 = np.random.randn(50)+1

trace0 = go.Box(
    y=y0
)
trace1 = go.Box(
    y=y1
)
data = [trace0, trace1]
py.iplot(data)
```


```{python}
```
